Analysis of split plot design using MCMCglmm

### 2018-04-09 README.Rmd
Update README.Rmd
Add badges and restructure to fit research compendium structure

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![Travis-CI Build Status](https://travis-ci.org/openplantpathology/rice_awd_pests.svg?branch=master)](https://travis-ci.org/openplantpathology/rice_awd_pests)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/phytopathology/rice_awd_pests?branch=master&svg=true)](https://ci.appveyor.com/project/phytopathology/rice_awd_pests)

# rice.awd.pests

The goal of _rice.awd.pests_ is to examine the effects of using alternate
wetting and drying irrigation techniques on sheath blight disease in rice
paddies.

## Study

**Objective 1:** Study the effects of different water and nitrogen managements
on sheath blight of rice

**Project PI:** Dr. Adam H Sparks

**Staff:**

  * Dr. Nancy P. Castilla (2016)

  * Mr. Michael Noel (2015)

******

### 1. SCHEDULE

IRRI Dry Season (2)

 * 2015 (December 2014 – April 2015)

 * 2016 (January 2016 - May 2016)

### 2. METHODOLOGY

**Experimental Design: Split-plot**

  * Main plot : Water management (2)

    * W1 – Alternate Wetting and Drying (AWD)

    * W2 – Flooded/Farmer’s Practice

  * Sub-plot : Nitrogen management (5)

    * 2015

      * N0 – 0 kg/ha

      * N1 – 100 kg/ha (farmer’s practice – based from J. Bigornia research)

      * N2 – 120 kg/ha (IRRI practice – based from L. Willocquet research)

    * 2016

      * N3 – 60 kg/ha

      * N4 - 180 kg/ha

  * Replication: Four (4)

  * Plot Sizes (2015)

    * Main plot size: 12m x 12m (144 sq m)

    * Sub-plot size: 5m x 5m (25 sq m)

    * Replication size: 12m x 24m (288 sq m)

    * Buffer: 1m per sub plot

    * Experiment size: 1,152 sq m

  * Plot Sizes (2016)

    * Main plot size

      * Block 2007: 21m x 20.5m (412.5 sq m)

      * Block 2008: 20.25m x 21.6m (437.4 sq m)

    * Sub-plot size

      * Block 2007: 21m x 10.25m (215.25 sq m)

      * Block 2008: 20.25m x 10.8m (218.7 sq m)

    * Replication size:

      * Block 2007: 42m x 20.5m (861 sq m)

      * Block 2008: 40.5m x 21.6m (874.8 sq m)

    * Buffer: 0.5m per sub plot

    * Experiment size: 3471.6 sq m

******

**Disease Rating Scale**

Disease was rated using an ordinal scale for field evaluation of sheath blight
(S Savary & N Castilla).

* **Rating Scale - Severity (%)**

    * 0 - 0,

    * 1 - 1%,

    * 2 - 5%,

    * 3 - 15%,

    * 4 - 50%,

    * 5 - >50%

******

**Nitrogen Management Rates**

<table width = "500">
<tr>
  <th rowspan = "2">Nitrogen<br>treatment</th>
  <th rowspan = "2">Year</th>
  <th rowspan = "2">Total N<br>(kg/ha)</th>
  <th colspan = "4">Application Amount (kg/ha)</th>
</tr>
<tr>
  <th>Basal</th>
  <th>Tillering</th>
  <th>Panicle<br>Intitiation</th>
</tr>
<tr>
  <td>N0</td>
  <td>2015</td>
  <td>0</td>
  <td>0</td>
  <td>0</td>
  <td>0</td>
</tr>
<tr>
  <td>N100</td>
  <td>2015</td>
  <td>100</td>
  <td>60</td>
  <td>20</td>
  <td>20</td>
</tr>
<tr>
  <td>N120</td>
  <td>2015</td>
  <td>120</td>
  <td>60</td>
  <td>30</td>
  <td>30</td>
</tr>
<tr>
  <td>N60</td>
  <td>2016</td>
  <td>60</td>
  <td>30</td>
  <td>30</td>
  <td>0</td>
</tr>
<tr>
  <td>N180</td>
  <td>2016</td>
  <td>180</td>
  <td>60</td>
  <td>60</td>
  <td>60</td>
</tr>
</table>

## Resources

### glmm split plot analysis

http://ase.tufts.edu/gsc/gradresources/guidetomixedmodelsinr/mixed%20model%20guide.html

http://www.stat.wisc.edu/courses/st572-larget/Spring2007/handouts16-4.pdf

http://www.maths.bath.ac.uk/~jjf23/mixchange/split.html

### Using MCMCglmm

http://www.maths.bath.ac.uk/~jjf23/mixchange/onewayanova.html

http://www.maths.bath.ac.uk/~jjf23/mixchange/split.html#mcmcglmm

https://github.com/tmalsburg/MCMCglmm-intro

### General GLMM in R

http://bbolker.github.io/mixedmodels-misc/glmmFAQ.html

### Citation

@@

### Licenses

Text and figures: [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

Code: See the [DESCRIPTION](DESCRIPTION) file

@@

### Contributions

We welcome contributions from everyone. Before you get started, please see our
[contributor guidelines](CONTRIBUTING.md). Please note that this project is
released with a [Contributor Code of Conduct](CONDUCT.md). By participating in
this project you agree to abide by its terms.

>>>>

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/openplantpathology/rice.awd.pests/commits/master)
[![minimal R version](https://img.shields.io/badge/R%3E%3D-`r as.character(getRversion())`-brightgreen.svg)](https://cran.r-project.org/)
[![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/) 
[![Travis-CI Build Status](https://travis-ci.org/openplantpathology/rice_awd_pests.svg?branch=master)](https://travis-ci.org/openplantpathology/rice_awd_pests)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/phytopathology/rice_awd_pests?branch=master&svg=true)](https://ci.appveyor.com/project/phytopathology/rice_awd_pests)

## Research compendium for a report on the effects of using alternate wetting and drying irrigation techniques on sheath blight disease in rice paddies.

### Compendium DOI:

http://dx.doi.org/xxxxxxx

The files at the URL above will generate the results as found in the
publication. The files hosted at github.com/openplantpathology/rice.awd.pests
are the development versions and may have changed since the report was
published.

### Authors of this repository:

Adam H. Sparks (adam.sparks@usq.edu.au)

Nancy P. Castilla (n.castilla@irri.org)

B. Ole Sander (b.sander@irri.org)

M. Noel

### Published in:

Sparks, A,  xxxxx

### Overview of contents

This repository is our research compendium for our analysis of the effects of
alternate wetting and drying irrigation technology on rice sheath blight
disease. The compendium contains all data, code, and text associated with the
publication. The `Rmd` files in the `analysis/paper/` directory contain details
of how all the analyses reported in the paper were conducted, as well as
instructions on how to rerun the analysis to reproduce the results. The
`data/` directory in the `analysis/` directory contains all the raw data.

### The supplementary files

The `analysis/` directory contains:

- the manuscript as submitted (in PDF format) and it's Rmd source file
- all the data files (in CSV file format, in the `data/` directory)
- supplementary information source files (in R markdown format) and executed versions
- all the figures that are included in the paper (in the `figures/` directory)

### The R package

This repository is organized as an R package. There are no actual R functions in
this package - all the R code is in the Rmd file. I simply used the R package
structure to help manage dependencies, to take advantage of continuous
integration for automated code testing, and so I didn't have to think too much
about how to organise the files.

To download the package source as you see it on GitHub, for offline browsing,
use this line at the shell prompt (assuming you have Git installed on your
computer):

```
git clone https://github.com/openplantpathology/rice_awd_pests.git
```

Once the download is complete, open the `rice.awd.pests.Rproj` in RStudio to
begin working with the package and compendium files.

The package has a number of dependencies on other R packages, and programs
outside of R. These are listed at the bottom of this README. Installing these
can be time-consuming and complicated, so we've done two things to simpify
access to the compendium. First is the packrat directory, which contains the
source code for all the packages we depend on. If all works well, these will be
installed on your computer when you open `rice_awd_.pests.Rproj` in RStudio.
Second is our Docker image that includes all the necessary software, code and
data to run our analysis. The Docker image may give a quicker entry point to the
project, and is more self-contained, so might save some fiddling with installing
things.

### The Docker image

A Docker image is a lightweight GNU/Linux virtual computer that can be run as a
piece of software on Windows and OSX (and other Linux systems). To capture the
complete computational environment used for this project we have a Dockerfile
that specifies how to make the Docker image that we developed this project in.
The Docker image includes all of the software dependencies needed to run the
code in this project, as well as the R package and other compendium files. To
launch the Docker image for this project, first, 
[install Docker](https://docs.docker.com/installation/) on your computer. At the
Docker prompt, enter:

    docker run -dp 8787:8787 adamhsparks/rice.awd.pests

This will start a server instance of RStudio. Then open your web browser at
localhost:8787 or or run `docker-machine ip default` in the shell to find the
correct IP address, and log in with rstudio/rstudio.

Once logged in, use the Files pane (bottom right) to navigate to `/` (the root
directory), then open the folder for this project, and open the `.Rproj` file
for this project. Once that's open, you'll see the `analysis/paper` directory in
the Files pane where you can find the R markdown document, and knit them to
produce the results in the paper. More information about using RStudio in Docker
is avaiable at the [Rocker](https://github.com/rocker-org) [wiki](https://github.com/rocker-org/rocker/wiki/Using-the-RStudio-image) pages.

We developed and tested the package on this Docker container, so this is the
only platform that We're confident it works on, and so recommend to anyone
wanting to use this package to generate the vignette, etc.

## Citation

@@

## Licenses

Manuscript: [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

Code: [MIT](http://opensource.org/licenses/MIT) year: 
`r format(Sys.Date(), "%Y")`, copyright holder: Adam Sparks

@@

## Dependencies

I used [RStudio](http://www.rstudio.com/products/rstudio/) on MacOS and Ubuntu
17.10. See the colophon section of the docx file in `analysis/paper` for a
full list of the packages that this project depends on.

## Contact

Adam Sparks, Associate Professor, Centre for Crop Health  
University of Southern Queensland, West St.  
Toowoomba, Qld 4350 Australia

(+61) 7.4631.1948   adam.sparks@usq.edu.au  
https://staffprofile.usq.edu.au/Profile/Adam-Sparks
